While I cannot possibly thank everyone that has been a part of this journey without forgetting a few, I will still make an attempt. 

I want to start by thanking my promoter, Chris. I am thankful for the freedom you gave me during the first half of my PhD, which allowed me to work on a variety of topics. I want to thank you for teaching me how to take criticism and think objectively about my work, especially in the latter half. I must thank you since it would have taken me a few more years to discover Westmalle Tripel if it were not for you.  

I want to thank my co-promoter, Anuradha. Some of our earlier discussions inspired me to think about glitches in the Einstein Telescope era. I am grateful for your support when working within the large collaborations and I also want to thank you for taking the time to hear me complain. 

I thank the administrators of the stoomboot computing cluster at Nikhef. I thank the GRASP secretaries, Monique and Marie-Thérèse, as well as the staff at the international service desk at Utrecht University. My master's thesis advisors, Soumen Roy and Anand Sengupta, deserve many thanks for getting me started on this topic. I appreciate the support from the collaborators in Leuven. Tjonnie, thank you for your insights on various occasions and for reminding me to slow down every time I felt like I was talking too slowly. I am thankful to Otto and the colleagues in Hong Kong.   

I would like to thank the colleagues at GRASP: Sumit, Mick, Fabian, Haris, Marteen, Luca, Olaf, Johanna, Gijs, Quirijn, Justus, Nick, Rick, Noor, Suzanne, Rene, Marcel, Marta, Alessandro, Thomas, and Raimond. Tomek, you have been a great friend. Thank you for helping me pay rent when I didn't yet have a Dutch bank account. I thank Anna for making the time to stay in touch, Stefano for the promises to stay in touch, Justin for the travels to conferences, and Bas for the good vibes. Thibeau and Peter, I look forward to working with you and to our next visit to the dim sum restaurant. Tom, I value our friendship and our collaboration. Marc, it was a pleasure sharing the office with you. 

I want to thank my friends Anam, Rajat, Melissa, Antoni, Anushree, Aayush, and Shridhar. I want to thank Marijn for facing all the headwind on our biking trips. My former landlords and landladies around Utrecht have also been a part of this adventure. If it weren't for your interventions, I might have completed my PhD a few months earlier.  

I am indebted to my parents, my brother, and my sister-in-law. I thank Suze's family for making me feel at home and Suze for her love.

